Equalizer 2.0\\Convergence of a Parallel Rendering Framework

### 2017-04-03 paperIEEE.tex
Flow and my CV
git-svn-id: https://dev.ifi.uzh.ch/vmml/svn/Equalizer16@30 db7c848b-7755-4df2-b763-8796797afa52

@@

In~\cite{EMP:09}, we presented the Equalizer parallel rendering framework. This
paper summarized the work in parallel rendering up to 2009. In the following we
will present the related work published since then. An extensive Programming and
User Guide provides in-depth documentation on using and programming
Equalizer~\cite{Eilemann:13}.

The concept of transparent OpenGL interception popularized by WireGL and
Chromium~\cite{HHNFAKK:02} has received little attention since 2009. While some
commercial implementations such as TechViz and MechDyne Conduit continue to
exist, on the research side only ClusterGL~\cite{NHM:11} has been
presented. ClusterGL employs the same approach as Chromium, but delivers a
significantly faster implementation of transparent OpenGL interception and
distribution for parallel rendering. CGLX~\cite{DK:11} tries to bring parallel
execution transparently to OpenGL applications, by emulating the GLUT API and
intercepting certain OpenGL calls. This approach works transparently for trivial
applications, but quickly requires the application developer to address the
complexities of a distributed application when mutable application state needs
to be synchronized across processes.

On the other hand, software for driving and interacting with tiled display walls
has received significant attention. Sage~\cite{Sage} and Sage2~\cite{Sage2} have
received significant attention. Sage is build entirely around the concept of a
shared framebuffer where all content windows are separate applications using
pixel streaming. It is no longer actively supported. Sage 2 is a complete,
browser-centric reimplementation where each application is a web application
distributed across browser instances. DisplayCluster~\cite{DC}, and its
continuation Tide, also implement the shared framebuffer concept of Sage, but
provide a few native content applications integrated into the display servers.

Equalizer itself has received significant attention to research and validate
parallel rendering performance, by using compression and region of interest
during compositing~\cite{MEP:10}, load-balancing resources for multi-display
installations~\cite{EEP:11}, asynchronous compositing and NUMA
optimizations~\cite{EBAHMP:12} as well as work queueing~\cite{SPEP:16}.

Various applications and frameworks have used Equalizer for new research in
visualization, such as the RTNeuron~\cite{HBBES:13}, an OpenGL renderer for
visualizing simulations of neucortical brain regions, Omegalib~\cite{Omegalib},
extending Equalizer for the hybrid 2D/3D environments such as the CAVE2,
visualization for remote sensing date~\cite{LK:09}, .

@@

hardware description

@@

\begin{figure}[ht]\center
  \includegraphics[width=.49\columnwidth]{images/spikes512}\hfil
  \includegraphics[width=.49\columnwidth]{images/spikes1280}
  \caption{\label{fVolumes}Spike frequencies of a three million neuron
    electrical simulation at 437x240x512 (51MB, left) and 1092x600x1280 (800MB,
    right) resolution}
\end{figure}

The spike frequency volumes aggregate the number of spikes which happened within
a given voxel over a given time. The absolute spike count is renormalized to an
unsigned byte value (0..255) during creation. Higher densities in the volume
represent higher spiking activity in the voxel. Consequently, the volume is more
sparsely populated and has a more even distribution of values at higher
resolution.

@@

\begin{IEEEbiography}{Michael Shell}
Biography text here.
\end{IEEEbiography}

>>>>

The rest of this paper is structured as follows: First we introduce new related
work since the introduction of Equalizer. The main body of this paper presents
new performance features, virtual reality algorithms, usability features to
build complex applications, main novel features of the underlying Collage
network library and a quick overview of the main Equalizer-based applications. A
result section presents new experiments not previously published, followed by
the discussion and conclusion.

@@

In~\cite{EMP:09}, we presented the Equalizer parallel rendering framework. This
paper summarized the work in parallel rendering up to 2009. An extensive
Programming and User Guide provides in-depth documentation on using and
programming Equalizer~\cite{Eilemann:13}. In the following we assume these two
publications and their references as the basis, and will present the related
work published since then.

The concept of transparent OpenGL interception popularized by WireGL and
Chromium~\cite{HHNFAKK:02} has received little attention since 2009. While some
commercial implementations such as TechViz and MechDyne Conduit continue to
exist, on the research side only ClusterGL~\cite{NHM:11} has been
presented. ClusterGL employs the same approach as Chromium, but delivers a
significantly faster implementation of transparent OpenGL interception and
distribution for parallel rendering. CGLX~\cite{DK:11} tries to bring parallel
execution transparently to OpenGL applications, by emulating the GLUT API and
intercepting certain OpenGL calls. In contrast to frameworks like Chromium, CGLX
follows the distribution application approach, whichworks transparently for
trivial applications, but quickly requires the application developer to address
the complexities of a distributed application when mutable application state
needs to be synchronized across processes.

On the other hand, software for driving and interacting with tiled display walls
has received significant attention. Sage~\cite{Sage} and Sage~2~\cite{Sage2}
have received significant attention. Sage is build entirely around the concept
of a shared framebuffer where all content windows are separate applications
using pixel streaming. It is no longer actively supported. Sage 2 is a complete,
browser-centric reimplementation where each application is a web application
distributed across browser instances. DisplayCluster~\cite{DC}, and its
continuation Tide~\cite{tide}, also implement the shared framebuffer concept of
Sage, but provide a few native content applications integrated into the display
servers.

Equalizer itself has received significant attention within the research
community. Various algorithms to improve the parallel rendering performance have
been proposed: compression and region of interest during
compositing~\cite{MEP:10}; load-balancing resources for multi-display
installations~\cite{EEP:11}; asynchronous compositing and NUMA
optimizations~\cite{EBAHMP:12}; as well as work queueing~\cite{SPEP:16}.

Various applications and frameworks have used Equalizer for new research in
visualization, such as the RTNeuron~\cite{HBBES:13}, an OpenGL renderer for
visualizing simulations of neucortical brain regions; Omegalib~\cite{Omegalib},
extending Equalizer for the hybrid 2D/3D environments such as the CAVE2;
visualization for remote sensing data~\cite{LK:09}, bino.

@@

This section presents new experiments, complementing the results of previous
publications~\cite{EP:07, EMP:09, MEP:10, EEP:11, EBAHMP:12, HBBES:13, deflect,
  SPEP:16}. The first part analyses data distribution performance, in particular
how the various optimizations in Collage effect complex operations in realistic
scenarios. The second part summarizes rendering performance over all
decomposition modes with a few representative workloads.

@@

\begin{figure}[ht]\center
  \includegraphics[width=.49\columnwidth]{images/spikes512}\hfil
  \includegraphics[width=.49\columnwidth]{images/spikes1280}
  \caption{\label{fVolumes}Volumes used for object distribution benchmarks:
    spike frequencies of a three million neuron electrical simulation at
    437x240x512 (51MB, left) and 1092x600x1280 (800MB, right) resolution}
\end{figure}

The spike frequency volumes aggregate the number of spikes which happened within
a given voxel over a given time. The absolute spike count is renormalized to an
unsigned byte value (0..255) during creation. Higher densities in the volume
represent higher spiking activity in the voxel. Consequently, the volume is more
sparsely populated and has a more even distribution of values at higher
resolutions.

@@

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{images/stefan}}]{Stefan Eilemann}
Stefan is the technical manager of the Visualization Team in the Blue Brain
Project. His work involves working towards large-scale visualization for
Exascale simulations; interactive integration of simulation, analysis and
visualization as well as flexible frameworks for data sharing and dynamic
allocation of heterogenous resources. He is the CEO and founder of Eyescale and
also the lead developer of the Equalizer parallel rendering framework. He has
worked in SGI's Advanced Graphics Division as the technical lead of OpenGL
Multipipe SDK - a toolkit for scalable, parallel visualization software on
shared memory supercomputers. He received his masters degree in Computer Science
from EPFL in 2015, and his Engineering Diploma in Computer Science in 1998.
\end{IEEEbiography}
